Fusion Moves for Correlation Clustering

### 2014-11-07 paper/egpaper_for_review.tex
update related work, relayout Figure EHC

\usepackage{eso-pic}
\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{booktabs}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
   We address the problem of partitioning a  graph
   into a previously unknown number of clusters.
   Among all partitions, the one with the minimal 
   sum of of cut edge weights is chosen. 
   This problem is known as correlation clustering 
   or multicut.
   We propose a general framework to find
   high quality approximate solutions for 
   this NP-hard problem based on a move making algorithm:
   We use candidate solutions from a proposal generator
   to iteratively improve the best observed solution similar
   to fusion moves.
   The proposed solver outperforms any other solver
   w.r.t. to any time performance.
   The solutions found by this solver are close
   to global optimal solutions w.r.t. energy
   and problem specific measurements as VI for
   image segmentation problems.

@@

Correlation clustering~\cite{Bansal-2004} also known as the multicut problem~\cite{chopra_1993_mp} 
is a basic primitive in computer vision~\cite{andres_2011_iccv,alush_2013_simbad} and data mining~\cite{Chierichetti-2014} as a first step
towards understanding an image or data in general. 
The correlation clustering problem is the following:
Given an undirected edge weighted graph, 
where positive weights (\emph{attractive}) encourage the adjacent nodes to be in the same cluster
and negative weights (\emph{repulsive}) encourage the nodes to stay in different cluster, 
find the clustering of nodes such that the sum over weights of edges between clusters is minimized.
Note that this problem formulation does not included a predefined number of clusters.

Despite the clear mathematical formulation and nice properties of correlation clustering
to perform a one-shot agglomeration clustering,
the main draw back of correlation clustering is scalability~\cite{nunez_iglesias_2013}
which arise from the NP-hardness of the problem.

For unsupervised image segmentation algorithms usually start with an over-segmentation
into superpixels (superregions), which are then clustered into ``perceptually meaningful''
regions by correlation clustering on sparse graphs.
Such an approach has been shown to yield
state-of-the-art results on the Berkeley Segmentation Database
\cite{andres_2011_iccv,yarkony_2012_eccv,alush_2013_simbad}.

Recently, Beier~\etal~\cite{Beier-2014} presented a promising 
approximative method called Cut, Clue \& Cut.
While this scales well for planar graphs, it does not for non planar graphs and has a bad any-time performance.
A further inherent drawback of this method is that the subproblems are not limited in their size.
%
Consequently, partition problems on large scale data, \eg
huge volume images in computational neuroscience~\cite{kroeger_2012_eccv}
or social networks~\cite{Leskovec-2010}, 
are not tractable, because the runtime explodes or the 
approximations are far away from being useful.

\textbf{Contribution:}
In the present work we present some novel approaches that are designed for large scale correlation clustering problems.
First, we define a novel energy based agglomerative clustering algorithm that monotonically increase the energy.
With this at hand we show how to improve the anytime performance of Cut, Clue \& Cut.
Second, we introduce cluster-fusion moves, which extend the original fusion moves~\cite{Lempinski-2010} 
used in supervised segmentation to the unsupervised case and give a polyhedral interpretation of this algorithm.
We propose two versatile proposal generators, and evaluate the proposed methods on existing and new benchmark problems.
That experiments show that we can improve the computation time by one to two magnitudes without worsening the segmentation 
quality significantly.

@@

\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two different regions.
%
A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
A segmentation of the graph $G$ can by either given by an 
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$.  
An edge labeling is only consistent if it contains no dangling edges~\cite{Kappes-2013-multicut}.
We denote the set of all consistent edge labelings by $P(G)\subset\{0,1\}^{|E|}$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G) = \textrm{conv}(P(G))$.

@@

Given a weighted graph $G=(V,E,w)$ we consider the problem of segmenting $G$ such that the costs
of the edges between distinct segments is minimized. This can be formulated in the node domain
by assigning each node $v$ a label $l_v \in \mathbb{N}$
\begin{align}
  l^* &= \argmin_{L \in \mathbb{N}^{|V|}} \sum_{ (i,j) \in E } w_{ij} \cdot [l_{i} \neq l_{j}], \label{eq:nodeproblem}
\end{align}  % y_{ij}^* &=& [l_{u} \neq l_{v}] 
or in the edge domain, by label each edge $e$ as cut $y_e=1$ or uncut $y_e=0$ 
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:edgeproblem}.%\\ 
\end{align}
As shown in~\cite{Kappes-2013-multicut} both problems are equivalent, but formulation \ref{eq:nodeproblem}
suffers from ambiguities in the formulation~\cite{Kappes-2011}.

@@

An other branch of research uses move making algorithm 
to optimize correlation clustering~\cite{bansal_2004_ml,beier_2014_cvpr,Kerninghan-1970}.
Starting with an initial segmentation, auxiliary problems are solved that 
strictly improve the segmentation.

@@

A more efficient move making method has been presented by Beier \etal~\cite{beier_2014_cvpr}.
Their Cut, Glue and Cut method iteratively re-optimizes the cuts between adjacent clusters of the current solution.
While this method scales better then  modified $\alpha-$expansion, it still faces three problems:
Firstly, the subproblems can be as large as the original problem,
secondly, the subproblems are max-cut problems on non-planar graphs that are NP-hard 
and even the used approximation has high polynomial complexity.
This limits the method, to be applicable to huge problems. The same holds for the 
method of Kerninghan and Lin~\cite{Kerninghan-1970}.

@@

For energy minimization problems fusion moves have become increasingly popular~\cite{Lepinsky-2010,kappes_2014_ws}.
For many large scale computer vision applications fusion moves lead to good approximations
with state of the art any time performance~\cite{kappes_2014_ws}.

@@

Outside computer vision greedy methods has been suggested for correlation clustering problems, see \cite{Elsner-2009} for an overview.
A common greedy approach is to randomly permute the nodes and than assign 
iteratively each node to an existing cluster or create a new cluster if costs cannot be decreased by assigning to a cluster.
Common assigning strategies are:
\emph{BEST}; assign to cluster which is linked by the most positive edge [Ng et al 2002 ACL],
\emph{FIRST}; Assign to the cluster which is first linked by a positive edge [Soon et al 2001 CL], and
\emph{VOTE}; Assign to cluster that minimizes objective function. [Elsner et al 08 ACL].

The PIVOT Algorithm [Ailon et al 2008 JACM] iterate over all nodes in random order.
If the node is not assigned it construct a cluster containing the node and all its 
unassigned positively linked neighbors.

Typically these algorithms started for many random permutations, 
and pick the clustering with best objective value.

A widely use post-processing method is  Best One Element Move (BOEM) [Gioniset al 2007 TKDD].
Start with an initial clustering one node is removed from a cluster and reassigned to an existing or new cluster such that the costs are minimized.
BEOM stops if no move can decrease the costs.

@@

@@

{\small
\bibliographystyle{ieee}
\bibliography{egbib}
}

>>>>

\usepackage{eso-pic}
\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{placeins}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
  \color{red} \textbf{TODO:}
   We address the problem of partitioning a  graph
   into a previously unknown number of clusters.
   Among all partitions, the one with the minimal 
   sum of of cut edge weights is chosen. 
   This problem is known as correlation clustering 
   or multicut.
   We propose a general framework to find
   high quality approximate solutions for 
   this NP-hard problem based on a move making algorithm:
   We use candidate solutions from a proposal generator
   to iteratively improve the best observed solution similar
   to fusion moves.
   The proposed solver outperforms any other solver
   w.r.t. to any time performance.
   The solutions found by this solver are close
   to global optimal solutions w.r.t. energy
   and problem specific measurements as VI for
   image segmentation problems.

@@

Correlation clustering~\cite{Bansal-2002} also known as the multicut problem~\cite{chopra_1993_mp} 
is a basic primitive in computer vision~\cite{andres_2011_iccv,kroeger_2012_eccv,yarkony_2012_eccv,alush_2013_simbad} and data mining~\cite{Chierichetti-2014,Arasu-2009,Sadikov-2010} as a first step
towards understanding an image or data in general. 
The correlation clustering problem is the following:
Given an undirected edge weighted graph, 
where positive weights (\emph{attractive}) encourage the adjacent nodes to be in the same cluster
and negative weights (\emph{repulsive}) encourage the nodes to stay in different cluster, 
find the clustering of nodes such that the sum over weights of edges between clusters is minimized.
Note that this problem formulation does not included a predefined number of clusters.

Despite the clear mathematical formulation and nice properties of correlation clustering
to perform a one-shot agglomeration clustering,
the main draw back of correlation clustering is scalability~\cite{nunez_iglesias_2013}
which arise from the NP-hardness of the problem~\cite{Bansal-2002}.

In signed social networks, where positive and negative edges encodes friend and foe relationships, respectively,
correlation clustering is a natural way to detect communities~\cite{Chierichetti-2014,Chen-2012}.
Correlation clustering can also been used to cluster query refinements in web search~\cite{Sadikov-2010}.
Because social and web-related networks are often very huge, heuristic methods, \eg the PIVOT-algorithm~\cite{Ailon-2008},
are very popular~\cite{Chierichetti-2014}.

In computer vision applications, unsupervised image segmentation algorithms usually start with an over-segmentation
into superpixels (superregions), which are then clustered into ``perceptually meaningful''
regions by correlation clustering on sparse graphs.
Such an approach has been shown to yield
state-of-the-art results on the Berkeley Segmentation Database
\cite{andres_2011_iccv,yarkony_2012_eccv,alush_2013_simbad}.

Recently, Beier~\etal~\cite{beier_2014_cvpr} presented a promising 
approximative method called Cut, Clue \& Cut.
While this scales well for planar graphs, it does not for non planar graphs and has a bad any-time performance.
A further inherent drawback of this method is that the subproblems are not limited in their size.
%
Consequently, partition problems on large scale data, \eg
huge volume images in computational neuroscience~\cite{kroeger_2012_eccv}
or social networks~\cite{Leskovec-2010}, 
are not tractable, because the runtime explodes or the 
approximations are far away from being useful.

In the present work we will suggest some alternative approaches to deal with large scale correlation clustering problems.
In the end we ill present a method that can be combined with any other correlation clustering method and 
allows to reduce the overall problem to a sequence of correlation clustering problems which are by magnitudes smaller.

\textbf{Contribution:}
In the present work we present some novel approaches that are designed for large scale correlation clustering problems.
First, we define a novel energy based agglomerative clustering algorithm that monotonically increase the energy.
With this at hand we show how to improve the anytime performance of Cut, Clue \& Cut.
Second, we introduce cluster-fusion moves, which extend the original fusion moves~\cite{Lempitsky-2010} 
used in supervised segmentation to the unsupervised case and give a polyhedral interpretation of this algorithm.
We propose two versatile proposal generators, and evaluate the proposed methods on existing and new benchmark problems.
That experiments show that we can improve the computation time by one to two magnitudes without worsening the segmentation 
quality significantly.

@@

\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two different regions.
%
A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
A segmentation of the graph $G$ can by either given by an 
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$.  
An edge labeling is only consistent if it contains no dangling edges~\cite{kappes_2013_arxiv}.
We denote the set of all consistent edge labelings by $P(G)\subset\{0,1\}^{|E|}$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G) = \textrm{conv}(P(G))$.

@@

Given a weighted graph $G=(V,E,w)$ we consider the problem of segmenting $G$ such that the costs
of the edges between distinct segments is minimized. This can be formulated in the node domain
by assigning each node $v$ a label $l_v \in \mathbb{N}$
\begin{align}
  l^* &= \argmin_{L \in \mathbb{N}^{|V|}} \sum_{ (i,j) \in E } w_{ij} \cdot [l_{i} \neq l_{j}], \label{eq:nodeproblem}
\end{align}  % y_{ij}^* &=& [l_{u} \neq l_{v}] 
or in the edge domain, by label each edge $e$ as cut $y_e=1$ or uncut $y_e=0$ 
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:edgeproblem}.%\\ 
\end{align}
As shown in~\cite{kappes_2013_arxiv} both problems are equivalent, but formulation \ref{eq:nodeproblem}
suffers from ambiguities in the formulation~\cite{kappes_2011_emmcvpr}.

@@

An other branch of research uses move making algorithm 
to optimize correlation clustering~\cite{bansal_2004_ml,beier_2014_cvpr,Kernighan-1970}.
Starting with an initial segmentation, auxiliary problems are solved that 
strictly improve the segmentation.

@@

A more efficient move making method has been presented by Beier \etal~\cite{beier_2014_cvpr}.
Their Cut, Glue and Cut method iteratively re-optimizes the cuts between adjacent clusters of the current solution.
While this method scales better then  modified $\alpha-$expansion, it still faces three problems:
Firstly, the subproblems can be as large as the original problem,
secondly, the subproblems are max-cut problems on non-planar graphs that are NP-hard 
and even the used approximation has high polynomial complexity.
This limits the method, to be applicable to huge problems. The same holds for the 
method of Kernighan and Lin~\cite{Kernighan-1970}.

@@

For energy minimization problems fusion moves have become increasingly popular~\cite{Lempitsky-2010,kappes_2014_ws}.
For many large scale computer vision applications fusion moves lead to good approximations
with state of the art any time performance~\cite{kappes_2014_ws}.

@@

Outside computer vision greedy methods has been suggested for correlation clustering problems, see \cite{Elsner-2009} for an overview.
A common greedy approach is to randomly permute the nodes and than assign 
iteratively each node to an existing cluster or create a new cluster if costs cannot be decreased by assigning to a cluster.
Common assigning strategies are:
\emph{BEST}; assign to cluster which is linked by the most positive edge~\cite{Ng-2002},
\emph{FIRST}; Assign to the cluster which is first linked by a positive edge~\cite{Soon-2001}, and
\emph{VOTE}; Assign to cluster that minimizes objective function~\cite{Elsner-2008}.
%
The PIVOT Algorithm~\cite{Ailon-2008} iterate over all nodes in random order.
If the node is not assigned it construct a cluster containing the node and all its 
unassigned positively linked neighbors.  
%
Typically these algorithms started for many random permutations, 
and pick the clustering with best objective value.

A widely use post-processing method is  Best One Element Move (BOEM)~\cite{Gionis-2007}.
Start with an initial clustering one node is removed from a cluster and reassigned to an existing or new cluster such that the costs are minimized.
BEOM stops if no move can decrease the costs.

@@

%\input{inputs/fig_hc_alg.tex}

@@

\FloatBarrier
{\small
\bibliographystyle{ieee}
\bibliography{egbib}
}
